---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r loadingData, echo=TRUE}
data <- read.csv(unz("activity.zip", "activity.csv"), header=T, colClasses = c("numeric", "Date", "numeric"))
print(summary(data))
```

## What is mean total number of steps taken per day?

```{r histogram, echo=TRUE}
dailySteps <- aggregate(data["steps"], by=data["date"], FUN = sum, na.rm = TRUE)
hist(dailySteps$steps, main="Total number of steps taken each day",
     xlab = "Number of steps")
mean <- mean(dailySteps$steps)
median <- median(dailySteps$steps)
```

The mean and median total number of steps taken per day are `r round(mean,1)` and `r format(median, scientific=FALSE)`, respectively.

## What is the average daily activity pattern?

```{r intervalAvg, echo=TRUE}
intervalsAvg <- aggregate(data["steps"], by=data["interval"], FUN = mean, na.rm = TRUE)
plot(intervalsAvg, type="l", main="Average daily activity pattern",
     xlab = "5-minute interval", ylab = "Number of steps")
maxInterval <- intervalsAvg$interval[which.max(intervalsAvg$steps)]
```
The 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps is `r maxInterval`.


## Imputing missing values

```{r missingValues, echo=TRUE}
nbNA <- sum(is.na(data$steps))
```
The total number of missing values is `r nbNA`.

Strategy applied for filling in all of the missing values in the dataset : a missing value is replaced by the mean for the corresponding 5-minute interval.

```{r fillingNA, echo=TRUE}
f <- function(x){
    
    if(!is.na(x["steps"])){
        y <- x["steps"]
    }
    else {
        y <- intervalsAvg$steps[intervalsAvg$interval==as.numeric(x[["interval"]])]
    }
    as.numeric(y)

}
data.filled <- data
data.filled$steps <- apply(data.filled,MARGIN=1,FUN=f)

dailySteps.filled <- aggregate(data.filled["steps"], by=data.filled["date"], FUN = sum, na.rm = TRUE)
hist(dailySteps.filled$steps, main="Total number of steps taken each day (filled dataset)",
     xlab = "Number of steps")
mean.filled <- mean(dailySteps.filled$steps)
median.filled <- median(dailySteps.filled$steps)

```
The mean and median total number of steps taken per day for the filled dataset are `r format(mean.filled, scientific=FALSE)` and `r format(median.filled, scientific=FALSE)`, respectively. These values differ from the estimates from the first part of the assignment. Imputing missing data resulted in the removal of the pronounced left-skew of the estimates from the first part.


## Are there differences in activity patterns between weekdays and weekends?
```{r weekendInfluence, echo=TRUE}
library(dplyr, warn.conflicts = FALSE)
# Suppress summarise info
options(dplyr.summarise.inform = FALSE)

library(lattice)
# Switch to english for weekdays()'s output.
invisible(Sys.setlocale("LC_TIME","English"))
mask.weekdays <- weekdays(data.filled$date) %in% c("Saturday","Sunday")
data.filled$dayType <- "weekday"
data.filled$dayType[mask.weekdays] <- "weekend"
data.filled$dayType <- as.factor(data.filled$dayType)
data.filled.group <- data.filled %>% group_by(dayType, interval) %>% summarise(steps = mean(steps))
xyplot(steps~interval | dayType, data=data.filled.group, layout=c(1,2), type="l",
       ylab = "Number of steps", xlab = "5-minute interval")
```
 

